Syntax Is All You Need: A Universal-Language Approach to Mutant Generation

### 2023-03-30 paper/main.tex
cheating a little

\begin{figure}[hbtp]
\centering
\caption{Mutation operator rule examples, using regular expressions (left) and the Comby language (right), based on parser parser combinators.}
\label{tab:rules}
  \begin{tabular}{l|l}
    \toprule
    \mc{1}{c|}{\textbf{Regexp}} & \mc{1}{c}{\textbf{Comby}}                          \\\midrule
    \mc{2}{c}{\textbf{Universal}}                         \\\midrule
    {\lstinline|< ==> >|} & {\lstinline|:[a]<:[b] ==> :[a]>:[b]|}   \\
    {\lstinline|< ==> ==|} & {\lstinline|:[a]<:[b] ==> :[a]==:[b]|} \\
    {\lstinline|< ==> <=|}  & {\lstinline|:[a]<:[b] ==> :[a]<=:[b]|} \\\midrule
\mc{2}{c}{\textbf{C-Like}} \\\midrule
 {\lstinline|else ==>|}  & {\lstinline|else ==>|} \\
 {\lstinline|if (\(.*\)) ==> if (!\1)|} &  {\lstinline|if (:[cond]) ==> if (!(:[cond]))|} \\
 {\lstinline|if (\(.*\)) ==> if (1==1)|} & {\lstinline|if (:[cond]) ==> if (1==1)|} \\
\toprule
\mc{2}{c}{\textbf{Language-Specific}} \\\midrule
 {\lstinline|synchronized ==>|} &  {\lstinline|synchronized ==>|} \\[0.5ex]
{\lstinline|(^\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf [\n]]|} \\
                                                         &\hspace{42pt}{\lstinline|==> :[s]pass:[lf]|} \\
 {\lstinline|any_of ==> all_of|} & {\lstinline|any_of ==> all_of|} \\
\bottomrule
\end{tabular}
\end{figure}

@@

\item Contiguous whitespace (e.g., spaces and newline) match contiguous whitespace in the source code. That is, match templates are sensitive to the \emph{presence} of whitespace, but \emph{not} the exact layout of spaces (i.e., the number of spaces may not correspond exactly between template and source code).

@@

\caption{Regex vs. Comby for C++ (top), Java (middle), and Rust (bottom). LOC describes the project files considered. Muts. is total number of mutants generated for those files. Invalid mutants are those that don't compile.}

@@

For Python, where TCE based on Python bytecode was possible, the
number of redundant (equivalent to the original bytecode, or
equivalent to the bytecode of an already-generated mutant) mutants was
overall similar for both Regex and Comby, though varied widely by file.

@@

\begin{table*}[hbtp]
    \centering
    \caption{Overview of Mutation Testing Tools. Our tool has significantly fewer lines of code than the single-language tools.}
    \label{tab:mutationtools}

    \begin{tabular}{|l|l|l|r|}
    \hline
    \textbf{Language} & \textbf{Tool} & \textbf{URL} & \textbf{LOC}  \\
    \hline
    \multirow{2}{*}{Python}  & Mutmut & https://github.com/boxed/mutmut & 3870  \\\cline{2-4}
        & cosmic-ray & https://github.com/sixty-north/cosmic-ray & 4599 \\ \hline
    \multirow{2}{*}{Java}  & PIT & https://github.com/hcoles/pitest & 59577  \\\cline{2-4}
        & LittleDarwin & https://github.com/aliparsai/LittleDarwin & 22359 \\ \hline
        Rust & Cargo-Mutants & https://github.com/sourcefrog/cargo-mutants & 7020 \\\hline
        C++ & Dextool & https://github.com/joakim-brannstrom/dextool & 38611 \\\hline
        ALL & Our Tool & removed for blinding & 2244 \\\hline
    \end{tabular}
    \end{table*}

@@

% ---------------- C++ table (UM vs Others)--------------------------
\begin{table}[htbp]
\centering
{\small
\caption{C++ (Our Implementation vs. Dextool). We cannot measure equivalent mutants for Dextool and thus omit the column.}
\label{tab:table_cpp2}
\begin{tabularx}{\columnwidth}{X|rrr|rr}
\toprule
        & \multicolumn{3}{c|}{\textbf{Comby}}  & \multicolumn{2}{c}{\textbf{Dextool}} \\ 
\textbf{Project} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} \\[1ex]\midrule
 ompl &  564 & 0.07 & 40\% & 278 & 0.08  \\
 libgraphqlparser  &  131 & 0.90 & 20\% & 52 & 0.96  \\
 rethinkdb\_rebirth &  492 & 0.10 & 20\% & 25 & 0.36 \\
 xoreos &  2280 & 0.01 & 10\% & 414 & 0.02  \\
 libxmljs &  790 & 0.64 & 10\% & 830 & 0.46 \\[0.5ex]
 tiny-differentiable-simulator &  770 & 0.04 & 0\% & 61 & 0.21  \\\midrule
\textbf{Mean}   & 837.83 & 0.29 & 16.67\% & 276.67 & 0.33\\
\textbf{Std.Dev.} & 745.78 & 0.37 & 13.66\%   & 311.56 & 0.38 \\\bottomrule
\end{tabularx}}
\end{table}

@@

\begin{table}[htbp]
\centering
\caption{Java (Our Implementation vs. PIT vs. LittleDarwin).
  Determining if PIT bytecode mutants are equivalent is extremely
  difficult to do with reasonable accuracy. LittleDarwin produces no equivalent mutants.}
\label{tab:table_java2}
\resizebox{\columnwidth}{!}{% again don't love it, but footnotesize is actually smaller than necessary
\begin{tabular}{l|rrr|rr|rr}
\toprule
                 & \mc{3}{c|}{\textbf{Comby}}   & \mc{2}{c|}{\textbf{PIT}} & \mc{2}{c}{\textbf{LittleDarwin}} \\
\textbf{Project} & \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} \\ \midrule

@@

\begin{table*}[htbp]
\centering
{
\caption{Python (Our Implementation vs. Mutmut vs. CosmicRay)}
\label{tab:table_python2}
\begin{tabular}{l|rrr|rrr|rrr}
\toprule
                 & \mc{3}{c|}{\textbf{Comby}}   & \mc{3}{c|}{\textbf{Mutmut}} & \mc{3}{c}{\textbf{CosmicRay}} \\
\textbf{Project} & \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv} 
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv} \\ \midrule
\midrule
keract & 1471 & 0.26 & 10\% & 482 & 0.72 & 10\% & 921 & 0.27 & 10\% \\
dtw & 866 & 0.55 & 10\% & 270 & 0.44 & 0\% & 713 & 0.46 & 20\% \\
notion-sdk-py & 196 & 0.60 & 20\% & 63 & 0.65 & 0\% & 52 & 0.44 & 0\% \\
atlassian-python-api & 333 & 0.62 & 40\% & 174 & 0.71 & 10\% & 106 & 0.74 & 10\% \\
ESD & 1712 & 0.07 & 10\% & 720 & 0.08 & 0\% & 1013 & 0.06 & 10\% \\
Fiber & 458 & 0.64 & 20\% & 222 & 0.57 & 40\% & 320 & 0.81 & 0\% \\\midrule
\textbf{Mean} &  839.33 & 0.46 & 18.00\% & 321.83 & 0.53 & 10.00\% & 520.93 & 0.46 & 8.33\% \\
\textbf{Std.Dev.} &  628.80 & 0.24 & 11.69\% & 239.04 & 0.24 & 15.49\% & 417.50 & 0.28 & 7.53\% \\\bottomrule        
\end{tabular}
}
\end{table*}

@@

  \begin{code}
 qsort :: [ Int] -> [ Int ]
 qsort [] = []
 qsort (x: xs ) = qsort l ++ [x] ++ qsort r
     where l = filter (< x) xs
           r = filter ( >= x) xs
\end{code}
\caption{Example from MuCheck Paper}
\label{fig:haskell}

\end{figure}

>>>>

\begin{figure}[hbtp]
\centering
\caption{\small Mutation operator rule examples, using regular expressions (left) and the Comby language (right), based on parser parser combinators.}
\label{tab:rules}
  \begin{tabular}{l|l}
    \toprule
    \mc{1}{c|}{\textbf{Regexp}} & \mc{1}{c}{\textbf{Comby}}                          \\\midrule
    \mc{2}{c}{\textbf{Universal}}                         \\\midrule
    {\lstinline|< ==> >|} & {\lstinline|:[a]<:[b] ==> :[a]>:[b]|}   \\
    {\lstinline|< ==> ==|} & {\lstinline|:[a]<:[b] ==> :[a]==:[b]|} \\
    {\lstinline|< ==> <=|}  & {\lstinline|:[a]<:[b] ==> :[a]<=:[b]|} \\\midrule
\mc{2}{c}{\textbf{C-Like}} \\\midrule
 {\lstinline|else ==>|}  & {\lstinline|else ==>|} \\
 {\lstinline|if (\(.*\)) ==> if (!\1)|} &  {\lstinline|if (:[cond]) ==> if (!(:[cond]))|} \\
 {\lstinline|if (\(.*\)) ==> if (1==1)|} & {\lstinline|if (:[cond]) ==> if (1==1)|} \\
\toprule
\mc{2}{c}{\textbf{Language-Specific}} \\\midrule
 {\lstinline|synchronized ==>|} &  {\lstinline|synchronized ==>|} \\[0.5ex]
{\lstinline|(^\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf [\n]]|} \\
                                                         &\hspace{42pt}{\lstinline|==> :[s]pass:[lf]|} \\
 {\lstinline|any_of ==> all_of|} & {\lstinline|any_of ==> all_of|} \\
\bottomrule
\end{tabular}
\end{figure}

@@

\item Contiguous whitespace (e.g., spaces, newline) match contiguous whitespace in the source code. That is, match templates are sensitive to the \emph{presence} of whitespace, but \emph{not} exact layout (i.e., the number of spaces may not correspond exactly between template and source code).

@@

\caption{\small Regex vs. Comby for C++ (top), Java (middle), and Rust (bottom). LOC describes the project files considered. Muts. is total number of mutants generated for those files. Invalid mutants are those that don't compile.}

@@

On Python, TCE on Python bytecode was possible. The
number of redundant (equivalent to the original bytecode, or
the bytecode of an already-generated mutant) mutants was
overall similar for both Regex and Comby, though varied widely by file.

@@

\begin{table*}[hbtp]
{\small
    \centering
    \caption{\small Overview of Mutation Testing Tools. Our tool has significantly fewer lines of code than the single-language tools.}
    \label{tab:mutationtools}

    \begin{tabular}{|l|l|l|r|}
    \hline
    \textbf{Language} & \textbf{Tool} & \textbf{URL} & \textbf{LOC}  \\
    \hline
    \multirow{2}{*}{Python}  & Mutmut & https://github.com/boxed/mutmut & 3870  \\\cline{2-4}
        & cosmic-ray & https://github.com/sixty-north/cosmic-ray & 4599 \\ \hline
    \multirow{2}{*}{Java}  & PIT & https://github.com/hcoles/pitest & 59577  \\\cline{2-4}
        & LittleDarwin & https://github.com/aliparsai/LittleDarwin & 22359 \\ \hline
        Rust & Cargo-Mutants & https://github.com/sourcefrog/cargo-mutants & 7020 \\\hline
        C++ & Dextool & https://github.com/joakim-brannstrom/dextool & 38611 \\\hline
        ALL & Our Tool & removed for blinding & 2244 \\\hline
    \end{tabular}}
    \end{table*}

@@

% ---------------- C++ table (UM vs Others)--------------------------
\begin{table}[htbp]
\centering
{\small
\caption{C++ (Our Implementation vs. Dextool). We cannot measure equivalent mutants for Dextool.}
\label{tab:table_cpp2}
\begin{tabularx}{\columnwidth}{X|rrr|rr}
\toprule
        & \multicolumn{3}{c|}{\textbf{Comby}}  & \multicolumn{2}{c}{\textbf{Dextool}} \\ 
\textbf{Project} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} \\[1ex]\midrule
 ompl &  564 & 0.07 & 40\% & 278 & 0.08  \\
 libgraphqlparser  &  131 & 0.90 & 20\% & 52 & 0.96  \\
 rethinkdb\_rebirth &  492 & 0.10 & 20\% & 25 & 0.36 \\
 xoreos &  2280 & 0.01 & 10\% & 414 & 0.02  \\
 libxmljs &  790 & 0.64 & 10\% & 830 & 0.46 \\[0.5ex]
 tiny-differentiable-simulator &  770 & 0.04 & 0\% & 61 & 0.21  \\\midrule
\textbf{Mean}   & 837.83 & 0.29 & 16.67\% & 276.67 & 0.33\\
\textbf{Std.Dev.} & 745.78 & 0.37 & 13.66\%   & 311.56 & 0.38 \\\bottomrule
\end{tabularx}}
\end{table}

@@

\begin{table}[htbp]
\centering
\caption{\small Java (Our Implementation vs. PIT vs. LittleDarwin).
  Accurately determining if PIT bytecode mutants are equivalent is intractable;
  LittleDarwin produces no equivalent mutants.}
\label{tab:table_java2}
\resizebox{\columnwidth}{!}{% again don't love it, but footnotesize is actually smaller than necessary
\begin{tabular}{l|rrr|rr|rr}
\toprule
                 & \mc{3}{c|}{\textbf{Comby}}   & \mc{2}{c|}{\textbf{PIT}} & \mc{2}{c}{\textbf{LittleDarwin}} \\
\textbf{Project} & \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} \\ \midrule

@@

\begin{table*}[htbp]
\centering
{
\caption{\small Python (Our Implementation vs. Mutmut vs. CosmicRay)}
\label{tab:table_python2}
\begin{tabular}{l|rrr|rrr|rrr}
\toprule
                 & \mc{3}{c|}{\textbf{Comby}}   & \mc{3}{c|}{\textbf{Mutmut}} & \mc{3}{c}{\textbf{CosmicRay}} \\
\textbf{Project} & \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv} 
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv} \\ \midrule
\midrule
keract & 1471 & 0.26 & 10\% & 482 & 0.72 & 10\% & 921 & 0.27 & 10\% \\
dtw & 866 & 0.55 & 10\% & 270 & 0.44 & 0\% & 713 & 0.46 & 20\% \\
notion-sdk-py & 196 & 0.60 & 20\% & 63 & 0.65 & 0\% & 52 & 0.44 & 0\% \\
atlassian-python-api & 333 & 0.62 & 40\% & 174 & 0.71 & 10\% & 106 & 0.74 & 10\% \\
ESD & 1712 & 0.07 & 10\% & 720 & 0.08 & 0\% & 1013 & 0.06 & 10\% \\
Fiber & 458 & 0.64 & 20\% & 222 & 0.57 & 40\% & 320 & 0.81 & 0\% \\\midrule
\textbf{Mean} &  839.33 & 0.46 & 18.00\% & 321.83 & 0.53 & 10.00\% & 520.93 & 0.46 & 8.33\% \\
\textbf{Std.Dev.} &  628.80 & 0.24 & 11.69\% & 239.04 & 0.24 & 15.49\% & 417.50 & 0.28 & 7.53\% \\\bottomrule        
\end{tabular}
}
\end{table*}

@@

\begin{lstlisting}
 qsort :: [ Int] -> [ Int ]
 qsort [] = []
 qsort (x: xs ) = qsort l ++ [x] ++ qsort r
     where l = filter (< x) xs
           r = filter ( >= x) xs
\end{lstlisting}
\caption{\small Example from MuCheck Paper~\cite{mucheck}}
\label{fig:haskell}
\end{figure}
